---
title: "Course Project : Practical Machine Learning"
author: "Joanne Antonetti"
date: "26 octobre 2014"
output: html_document
---



### Summary of the problem
> Question: What is the class (A, B, C, D, E) of a sample?

Input Data: 
- 19622 samples
- 159 predictors
- 5 classes



### 1) Data Preparation
#### Data loading
```{r}
training <- read.csv("pml-training.csv", na.strings= c(" ", "NA",""))
testing <- read.csv("pml-testing.csv", na.strings= c("NA",""," "))
```

```{r}
dim(training) 
```

```{r}
dim(testing) 
```


#### Data Cleaning of the testing and training set
We remove the variables with the NA values.
```{r}
training_clean <- training[ , colSums(is.na(training)) == 0]
testing_clean <- testing[ , colSums(is.na(testing)) == 0]
```

```{r}
dim(training_clean) 
```

```{r}
dim(testing_clean)
```
  
We remove the variables which are not relevant for this analysis.
```{r}
training_clean <- training_clean[,8:60]
testing_clean <- testing_clean[,8:60]
```

```{r}
dim(training_clean) 
```

```{r}
dim(testing_clean) 
```



### 2) The Model  
#### Data splitting
We split data into training and testing subset for cross validation. 
```{r}
library(caret)
library(lattice)
library(ggplot2)
set.seed(32323)
inTrain = createDataPartition(y=training_clean$classe, p=0.75, list=FALSE)
training_train = training_clean[inTrain,]
training_test = training_clean[-inTrain,]
```  


#### Training
We choose a type of model on the training set.
```{r}
library(randomForest)
set.seed(121212)
modFit <- randomForest(classe~.,data=training_train,importance=TRUE)
print(modFit)
``` 


#### Prediction on testing to evaluate the model : Out of Sample Error
The expected **Out of Sample Error** is estimated by the error rate on the test subset, that is  *1 - Accuracy = 0.0047*
```{r}
predictions <- predict(modFit, newdata=training_test)
confMatrix <- confusionMatrix(predictions, training_test$classe)
print(confMatrix)
``` 


#### Matrix representing the importance of the predictors 
```{r}
varImpPlot(modFit, main="Variable Importance")
``` 



###  3) Prediction on the testing set
#### Prediction
```{r}
answers <- predict (modFit, testing_clean)
print(answers)
``` 


#### Generating Answers Files to Submit for Assignment
```{r}
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(answers)
```

